Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-28 WSSSPE1_report/WSSSPE1_summary_report.tex
adding career path discussion

Some examples of such software discussed as part of the workshop are,
\emph{DUNE}, \emph{Kitware}, \emph{PETSc}, \emph{MINRES-QLP}, \emph{FASTMath}.
Most of the above efforts are focused on one scientific and/or applied
mathematics domain. However, sometimes such projects grow beyond the initial
vision of developers. One such example is \emph{Kitware}, which while being a
software product specializing in the scientific process, has a core focus of
developing communities around software processes. An instance of this process
is the development of the \emph{CMake} build utility, which started out as a
building tool for \emph{ITK} but grew to become a generic build utility for C++
projects.
%

%Summary of above discussed cases: lessons, trends, commonalities, is there an ideal sustainability scenario?
It appears that the career paths for utility software developers are broader
compared to the developers of niche software. However, as projects as a whole,
the small-scale, specific software have a lower risk and cost in terms of
sustaining as compared to utility software. The development of scientific
software requires more scientific skills compared with those of utility
software. This diminishes the need of specialist programmers and practices in
many cases where the bulk of development is carried out by a domain scientist.

>>>>

Some examples of such software discussed as part of the workshop are,
\emph{DUNE}, \emph{R/qtl}, \emph{Kitware}, \emph{PETSc}, \emph{MINRES-QLP},
\emph{FASTMath}.  Most of the above efforts are focused on one scientific
and/or applied mathematics domain. However, sometimes such projects grow beyond
the initial vision of developers. One such example is \emph{Kitware}, which
while being a software product specializing in the scientific process, has a
core focus of developing communities around software processes. An instance of
this process is the development of the \emph{CMake} build utility, which
started out as a building tool for \emph{ITK} but grew to become a generic
build utility for C++ projects. Similarly, \emph{PETSc} is growing towards
becoming a general purpose utility system usable for solving variety of
scientific problems.
%

%Summary of above discussed cases: lessons, trends, commonalities, is there an ideal sustainability scenario?
In order to investigate the much discussed issue of career paths for developers
of scientific software, we conducted a small survey of the use-cases discussed
above with respect to the current status of its core developers. We looked at
the workshop submissions and the original cited papers for the respective
software. We looked at the webpages and source codes of the earliest versions
where available to find the original authors. we find that in a majority of
scientific software more than one original developers are still around, in some
cases for more than 10 years. Whereas, in the case of utility software
projects, it appears that at least one original developer is still around often
in a part-time role. In other cases, it appears that the original developers
have went on to assume a larger role, again with the same project.

From this survey, it appears that the career paths for utility software
developers are broader compared to the developers of niche software. However,
as projects as a whole, the small-scale, specific software have a lower risk
and cost in terms of sustaining as compared to utility software. Often the
sustainability of such projects are achieved by the fact that the core
developer or team heavily utilizes the software for their science, eg.
\emph{R/qtl, PETSc}. The development of scientific software requires more
scientific skills compared with those of utility software. This diminishes the
need of specialist programmers and practices in many cases where the bulk of
development is carried out by a domain scientist.
